### 2011-09-20 doc/theses/mikula/analysis_of_recent_approaches.tex
Master thesis: In progress.

\section{Discovering XML Keys and Foreign Keys in Queries}
This method is described in paper \cite{Necasky:2009:DXK:1529282.1529414}. It tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries. Just the queries are utilized in inference, no XML data are used. The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} \todo{aj unique?} constructs.

\subsection{Keys and foreign keys}
The authors introduce a formalism that formalizes XML Schema keys.

\todo[inline]{Definovat, co pouzivame: XPath (aj s verziou), prikady, osy, lomitko, ...}

\begin{define}[Key]
A key is a construct $$(C, P, \{L\})$$ where $C$, $P$, $L$ are XPath paths without predicates that use only \texttt{child} and \texttt{descendant} axes.
$C$ can be omitted. $(P, \{L\})$ is then equivalent to $(/, P, \{L\})$ and is called \emph{global key}. Otherwise it is called \emph{local key}.
\end{define}

The key specifies the following condition. Let $c$ be an element targeted by $C$ and $p$ and $p’$ be two elements targeted by $P$ from $c$. If the value targeted by $L$ from $p$ equals to the value targeted by $L$ from $p’$, then $p$ and $p’$ are the same elements. In other words, no two different elements targeted by $P$ from $c$ can have the same value of $L$.

\begin{define}[Foreign key]
A foreign key is a construct $$(C, (P_1, \{L_1\}) \Rightarrow (P_2, \{L_2\}))$$ where $(C, P_2, \{L_2\})$ is a key, $P_1$, $L_1$ are XPath paths without predicates that use only \texttt{child} and \texttt{descendant} axes.
$C$ can be omitted as in the case of keys.
\end{define}

\todo[inline]{Potřebujeme tyto definice k něčemu dále?}

Let $c$ be an element targeted by $C$ and $p_1$ be a element targeted by $P_1$ from $c$. The foreign key specifies that there is an element $p_2$ targeted by $P_2$ from $c$ such that the value targeted by $L_1$ from $p_1$ equals to the value targeted by $L_2$ from $p_2$. In other words, each element targeted by $P_1$ from $c$ refers to an element targeted by $P_2$ from $c$ via the pair $L_1$ and $L_2$.

\subsection{Key and Foreign Key Discovery}
\subsubsection{Joins in queries}
To discover keys and foreign keys, the described method utilizes element/element joins.

Assume a query $Q$ that joins a sequence of elements $S_1$ targeted by a path $P_1$ with a sequence of elements $S_2$ targeted by a path $P_2$ on a condition $L_1 = L_2$. It means that $Q$ joins an element $e_1$ from $S_1$ with an element $e_2$ from $S_2$ if $e_1/L_1$ equals to $e_2/L_2$.

\todo[inline]{priklad?}
\todo[inline]{Určitě. A hlavně je tento popis mnohem detailnější, přesnější a formaálnější než ty předchozí. Mělo by to být všechno v jednodném stylu.}

Assuming that each join is done via a key/foreign key pair, it can be inferred from $Q$ that $L_1$ is a key for elements in $S_1$ or $L_2$ is a key for elements in $S_2$ and the other is a foreign key referencing the key.

The authors introduce two following observations:
\begin{enumerate}
\renewcommand{\theenumi}{(O\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item Assume that an element $e_1$ from $S_1$ can be joined with more elements from $S_2$. It means that there can be more different elements in $S_2$ having their value of $L_2$ equal to $e_1/L_1$. In other words, there can be more different elements in $S_2$ having the same value of $L_2$. Therefore, we infer that $L_2$ can not be a key of the elements in $S_2$. Moreover, because we suppose that one of $L_1$ and $L_2$ is a key and $L_2$ can not be a key, we infer that $L_1$ is a key of the elements in $S_1$. Consequently, we infer that $L_2$ is a foreign key referring $L_1$.
\item Vice versa, assume that $e_1$ can be joined with maximally one element from $S_2$. It means that there is maximally one element in $S_2$ having its value of $L_2$ equal to $e_1/L_1$. In that case we suppose that each element in $S_2$ has a unique value of $L_2$. Therefore, we infer that $L_2$ is a key of the elements in $S_2$ and consequently that $L_1$ is a foreign key referring $L_2$. We cannot infer whether $L_1$ is a key or not.
\end{enumerate}

\subsubsection{Join patterns}
For a certain join, the decision for one of the cases (O1) and (O2) is made by the form of the join. The query is searched for so-called \emph{join patters}. These are \texttt{for} \emph{join pattern} and \texttt{let} \emph{join pattern} and they look as follows (the first is \texttt{for} \emph{join pattern}, the second is \texttt{let} \emph{join pattern}):

\todo[inline]{spravne kreslit indexy pri pismenkach}
\begin{verbatim}
01  for e1 in P1               for $e1 in P1
02  return                     return
03    for $e2 in                 let $e2 :=
        P2[L2 = $e1/L1]            P2[L2 = $e1/L1]
04    return CR                  return CR
\end{verbatim}

$P_1$, $P_2$, $L_1$ and $L_2$ in the patterns are XPath paths without predicates. Both patterns differ only at line \texttt{03} where the former applies \texttt{for} while the other applies \texttt{let}. Line \texttt{01} is called \emph{declaration clause}. Line \texttt{03} is called \emph{join clause}. The path $P_1$ is called source path, $P_2$ \emph{target path}, and the condition $L_2 = \$e_1/L_1$ \emph{join condition}. $C_R$ is called \emph{return clause}. All paths $P_1^R$, ..., $P_k^R$ in $C_R$
starting with $\$e2$ are called \emph{target return paths}.

Each pattern occurrence is marked as \emph{repeating} or \emph{non-repeating}. An occurrence marked as \emph{repeating} means that each element targeted by $P_1$ can be joined with more than one elements targeted by $P_2$. In contrast, a pattern occurrence marked as \emph{non-repeating} means that each element targeted by $P_1$ can be joined with zero or one element targeted by $P_2$. In other words, the former means that the observation (O1) can be applied while the latter means that (O2) can be applied to infer keys and foreign keys.

Assume a pattern occurrence $\pi$. The decision is made on the basis of the following rules (R1) - (R5). Only one rule can be applied. Process of decision starts with (R1). If (Ri) can not be applied, it tries (Ri+1). If any of the following rules (R1) - (R3) can be applied, $\pi$ is marked as \emph{repeating}. Each pattern occurence is also assigned with its weight determining how sure the method is about the inferred statement.

\begin{enumerate}
\renewcommand{\theenumi}{(R\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item $\pi$ is an occurrence of the \texttt{for} pattern (weight: 1).
\item Aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path $P_i^R$ in $\pi$ (weight: 1).
\item Aggregation function \texttt{count} is applied on a \emph{return path} $P_i^R$ in $\pi$ (weight: 0.75).
\end{enumerate}

If (R1) - (R3) can not be applied, $\pi$ is marked as \emph{non-repeating}. The weight is assigned according to (R4) and (R5) listed bellow where $k$ denotes the number of target \emph{return paths} in $\pi$.

\begin{enumerate}
\renewcommand{\theenumi}{(R\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\setcounter{enumi}{4}
\item $k > 1$ (weight: 1)
\item$ k <= 1$ (weight 0.5)
\end{enumerate}

\subsubsection{Key and foreign key inference}
Let $w$ be the weight assigned to a pattern occurrence $\pi$. If $\pi$ is marked as \emph{repeating}, (O1) is applied and the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(P_2^{\downarrow}, \{L_2\})$ is not satisfied
\item $(P_1^{\downarrow}, \{L_1\})$ is satisfied
\item $(P_2^{\downarrow}, \{L_2\}) \Rightarrow (P_1^{\downarrow}, \{L_1\})$ is satisfied
\end{itemize}
where for a path $P$, $P^{\downarrow}$ denotes:
\begin{itemize}
\item $P$, if $P$ does not start with a variable
\item $P^{\$e\downarrow}/P'$ (or $P^{\$e\downarrow}//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) where $P^{\$e}$ is a path used for the declaration of the variable $\$e$ in the query
\end{itemize}

If $\pi$ is marked as \emph{non-repeating}, (O2) is applied and the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(P_2^{\downarrow}, \{L_2\})$ is satisfied
\item $(P_1^{\downarrow}, \{L_1\}) \Rightarrow (P_2^{\downarrow}, \{L_2\})$ is satisfied
\end{itemize}

If the \emph{source} and \emph{target path} in $\pi$ have a common context determined by a path $C$, keys and foreign keys are inferred as follows. If $\pi$ is marked as \emph{repeating}, the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(C^{\downarrow}, P_2^{\downarrow C}, \{L_2\})$ is not satisfied
\item $(C^{\downarrow}, P_1^{\downarrow C}, \{L_1\})$ is satisfied
\item $(C^{\downarrow}, (P_2^{\downarrow C}, \{L_2\}) \Rightarrow (P_1^{\downarrow C}, \{L_1\}))$ is satisfied
\end{itemize}
where for paths $C$ and $P$, $P^{\downarrow C}$ denotes:
\begin{itemize}
\item $P$, if $P$ does not start with a variable
\item $P'$ (or $.//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) and $\$e$ is declared by $C$
\item $P^{\$e\downarrow C}/P'$ (or $P^{\$e\downarrow C}//P'$), if $P$ is $\$e/P'$ (or $\$e//P'$, respectively) and $\$e$ is not declared by $C$
\end{itemize}

If $\pi$ is marked as \emph{non-repeating}, the following statements with weight $w$ are inferred:
\begin{itemize}
\item $(C^{\downarrow}, P_2^{\downarrow C}, \{L_2\})$ is satisfied
\item $(C^{\downarrow}, (P_1^{\downarrow C}, \{L_1\}) \Rightarrow (P_2^{\downarrow C}, \{L_2\}))$ is satisfied
\end{itemize}

\todo[inline]{Problém je, že tady ste v podstatě opsal ten článek. To je špatně. Analýza ex. přístupů by měla podat přehled toho, co se dnes v dané oblasti umí. Každý přístup by měl být popsán zhruba stejným stylem, stejně podrobně a s jednotným značením. Tj. ne zkopírovat různé odstavce z různých článků, z něčeho 20 řádků a z něčeho 6 stran. To není analýza ale plagiátorství.}

\section{Scoring function}
It can happen that some of the assumptions the method is based on are not met. Assuming a set of queries, a key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key. \todo[inline]{priklad?}

Therefore, the authors introduce a scoring function. If a new key $K$ is going to be inferred, it is assigned with an initial score 0. Each inferred positive statement about $K$ increases the score of $K$ by the weight of the statement. Respectively, each negative statement about $K$ decreases the score by the respective weight. The resulting score therefore summarizes the weights of all inferred statements about $K$. A positive score means that $K$ is probably satisfied while negative means that $K$ is probably not satisfied. The higher the absolute value of the score is, the higher the probability is.

Let $K_1$, ..., $K_n$ be the inferred keys. Let $S_i$ be the score of $K_i$ and $N_i$ be the number of the inferred statements about $K_i$. The precision of the scoring is further enhanced on the base of the following observation. Assume a key $K_i = (C, P, \{L\})$ and $K_j = (C' , P, \{L\})$ where the path $C$ targets ancestors of the elements targeted by $C'$, i.e. the context specified by $C$ covers the context specified by $C'$. In that case it is also said that $K_i$ \emph{covers} $K_j$. It can be easily seen that if $K_i$ is satisfied, $K_j$ must be satisfied as well. On the other hand, if $K_j$ is not satisfied, $K_i$ can not be satisfied too. Therefore, if the score $S_i$ of $K_i$ is positive (i.e. $K_i$ is satisfied), $S_j$ is incremented with $S_i$ and $N_j$ with $N_i$ (i.e. $K_j$ is satisfied as well). Conversely, if the score $S_j$ of $K_j$ is negative (i.e. $K_j$ is not satisfied), $S_i$ is incremented with $S_j$ and $N_i$ with $N_j$ (i.e. $K_i$ can not be satisfied too).

Finally, the scores of the inferred keys are normalized to be comparable with each other. The normalized scores are from the range $\langle -1, 1\rangle$. The normalization takes into account not only the scores summarizing the weights of the statements about keys but also the number of the statements. The normalized score is computed as follows. Let $S^{max}$ be the maximum from $|S_1|$, ..., $|S_n|$ and $N^{max}$ be the maximum from $N_1$, ..., $N_n$. The normalized score $S_i^{norm}$ of $K_i$ is computed as follows:
$$S_i^{norm} = {S_i \over S^{max}} * (1 - {N^{max} - N_i \over \sum _{i=1}^n N_i})$$

\section{Conclusion}
The output of the method is a list of scored keys and for each key a list foreign keys referencing the key. The score of a key can be negative or positive. A negative score means that the key is not specified by the XML documents while positive means that it is satisfied. The absolute value of the score means how sure the method is about it.

>>>>

\section{Discovering XML Keys and Foreign Keys in Queries}
This method is described in paper \cite{Necasky:2009:DXK:1529282.1529414}. It tries to improve automatic XML schema inference by discovering keys and foreign keys from a set of XQuery queries. Just the queries are utilized in inference, no XML data are used. The output of this method is a set of keys and foreign keys that can be captured using XML Schema \textbf{key}, \textbf{keyref} and \textbf{unique} constructs.

\subsection{Assumptions and Observations}
To discover keys and foreign keys, the method utilizes element/element joins. Assume a query $Q$ that joins a sequence of elements $S_1$ targeted by a path $P_1$ with a sequence of elements $S_2$ targeted by a path $P_2$ on a condition $L_1 = L_2$.
\todo[inline]{priklad}
The method is based on assumption that each join is done via key/foreign key pair. It means it is supposed that $L_1$ is a key of the elements $S_1$ and $L_2$ is it's respective foreign key or vice versa.

The author describe two possible cases:

\begin{itemize}
\renewcommand{\theenumi}{(O\arabic{enumi})}
\renewcommand{\labelenumi}{\theenumi}
\item $L_1$ is a key of the elements $S_1$, $L_2$ is a respective foreign key and it itself is not a key of the elements $S_2$.
\item $L_2$ is a key of the elements $S_2$, $L_1$ is a respective foreign key and it cannot be decided whether $L_1$ is a key of the elements $S_1$ or not.
\end{itemize}

\subsection{Join Patterns and Key Inference}
For a certain join, the decision for one of the cases (O1) and (O2) is made by the form of the join. The query is searched for so-called \emph{join patters}. These are \emph{for join pattern} and \emph{let join pattern} and they are as follows (the first is \emph{for join pattern}, \emph{let join pattern}):

\todo[inline]{spravne kreslit indexy pri pismenkach}
\begin{verbatim}
for $e1 in P1              for $e1 in P1
return                     return
    for $e2 in                 let $e2 :=
        P2[L2 = $e1/L1]            P2[L2 = $e1/L1]
    return CR                  return CR
\end{verbatim}

Each occurrence of a join pattern is classified by application of the following rules it this specific order. The first satisfied ruled is applied. The occurrence is also assigned with a weight determining how sure the method is about the inferred statement.

The pattern occurrence is considered of case (O1) if it is the for join pattern (weight: 1), aggregation function \texttt{avg}, \texttt{min}, \texttt{max} or \texttt{sum} is applied on a return path (weight: 1), aggregation function \texttt{count} is applied on a return path. \todo {definovat return path?}

Else, the pattern occurrence is considered of case (O2) and the assigned weight depends on the number of target return paths. If the number is greater than one, the weight is one (weight: 1), else (the number is less or equal one) the weight is one half (weight: 0.5).

\subsection{Summarization of the Results}

The assumption the method is based on may not be fulfilled for every join in a supplied set of queries. A key $K$ may be inferred from some query and processing of another query may result to denial of $K$ as a key. \todo[inline]{priklad?}

Therefore, the authors introduce a scoring function to summarize the positive and the negative statements about keys using the assigned weights. The value of the score expresses the probalitity that a respective key statement is satisfied. Finally, the scores of the inferred keys are normalized to be comparable with each other.

\subsection{Conclusion}
The output of the method is a list of scored keys and for each key a list foreign keys referencing the key. Since the method deals only with the inference of keys, it is not an overall method of the XML schema inference. It is meant to be used in a collaboration with other schema-inferring methods to refine their results.
